---
title: SjS MSG [scRNA-seq]
author: 
  name: Jun Inamo
  email: juninamo@keio.jp
  affiliation: Division of Rheumatology, Department of Internal Medicine, Keio University School of Medicine, Tokyo, Japan
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  BiocStyle::html_document:
    toc_float: true
---



```{r, echo=FALSE, results="hide", message=FALSE}
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
# BiocManager::install("BiocStyle")
library(BiocStyle)
```

```{r}
source("utils.R")
```


```{r}

data_type = "GEX"

successful_gex_list <- list()
successful_gex_list_mat <- list()

for (sample_id in sample_ids) {
  tryCatch({
    gex.data <- Read_CellBender_h5_Mat(paste0("/path/to/",sample_id,"/cellbender_feature_bc_matrix_filtered.h5"))
    successful_gex_list_mat[[sample_id]] <- gex.data
    gex <- CreateSeuratObject(counts = gex.data, 
                              project = paste0(data_type,"_",sample_id), min.cells = 3, min.features = 200)
    
    successful_gex_list[[sample_id]] <- gex
  }, error = function(e) {
    print(paste("Error at sample_id =", sample_id))
  })
}
print("Success sample_ids; ")
print(names(successful_gex_list))

cell_bender_merged <- Merge_Sparse_Data_All(matrix_list = successful_gex_list_mat, 
            add_cell_ids = names(successful_gex_list_mat))

gex <- Merge_Seurat_List(successful_gex_list, 
                         add.cell.ids = names(successful_gex_list),
                         merge.data = TRUE)

```



```{r}

# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
gex[["percent.mt"]] <- PercentageFeatureSet(gex, pattern = "^MT-")

```


```{r}

gex <- subset(gex, subset = nFeature_RNA > 200 & nFeature_RNA < 5000 & percent.mt < 5)
rb.genes <- rownames(gex)[grep("^RP[SL]",rownames(gex))]
percent.ribo <- colSums(gex[rb.genes,])/Matrix::colSums(gex)*100
gex <- AddMetaData(gex, percent.ribo, col.name = "percent.ribo")

```


## Load metadata

```{r}

metaInfo = dplyr::left_join(gex@meta.data,
                            meta %>%
                              dplyr::mutate(orig.ident = paste0(data_type,"_",sample)),
                            by="orig.ident")

gex@meta.data$Age = metaInfo$Age
gex@meta.data$Sex = metaInfo$Sex
gex@meta.data$Disease = metaInfo$Disease

```



## Normalizing the data

```{r}
gex <- NormalizeData(gex, normalization.method = "LogNormalize")
gex <- NormalizeData(gex)
```

## Identification of highly variable features (feature selection)

```{r}
gex <- FindVariableFeatures(gex, selection.method = "vst", nfeatures = 2000)

```

## Scaling the data
```{r}
gex <- ScaleData(gex)
```

## Perform PCA

```{r}
set.seed(1234)
gex <- RunPCA(gex, features = VariableFeatures(object = gex))
```

## batch correction

```{r}
set.seed(1234)
gex <- gex %>% 
  RunHarmony(c("batch"), plot_convergence = TRUE)
```



## Cluster cells


```{r}
# use best number of hPCs and min.dist for UMAP
dim=20
min.dist=0.3
resolution=0.1
gex <- FindNeighbors(gex, reduction = "harmony", 
                     dims = 1:dim,
                     k.param = 30,
                     verbose = FALSE
)
gex <- RunUMAP(gex, reduction = "harmony", 
               dims = 1:dim,
               n.neighbors = 30L,
               min.dist = min.dist,
               verbose = FALSE)
gex <- FindClusters(gex, reduction = "harmony", 
                    dims = 1:dim, 
                    resolution = resolution,
                    verbose = FALSE)
```


## filter our cells in clusters with < 30 cells

```{r}
min_cell_cluster = 30

gex = subset(gex, 
             idents = names(table(Idents(gex)))[table(Idents(gex))>min_cell_cluster])

```


## Association test by cluster-free and cluster-based approach

### MiloR

```{r}
disease_group = "All"
disease_label = "AllSS"
control_group = c("nonSS")

milo <- Milo(as.SingleCellExperiment(gex))
milo <- buildGraph(milo, 
                   reduced.dim = "HARMONY",
                   k = 30, d = 30)
milo <- makeNhoods(milo, prop = 0.1, k = 30, d=30, 
                   refined = TRUE, reduced_dims = "HARMONY")
milo <- countCells(milo, meta.data = data.frame(colData(milo)), samples="orig.ident")

milo_design <- data.frame(colData(milo))[,c("orig.ident", "Disease",
                                            "Age","Sex",
                                            "batch")]
milo_design <- distinct(milo_design)
rownames(milo_design) <- milo_design$orig.ident
milo_design <- milo_design[colnames(nhoodCounts(milo)), , drop=FALSE]

contrast.1 <- paste0("Disease",disease_label, " - DiseasenonSS") 
da_results <- testNhoods(milo,
                         design = ~ 0 + Age + Sex + batch + Disease, 
                         design.df = milo_design, 
                         model.contrasts = contrast.1, 
                         fdr.weighting="graph-overlap")

da_results <- annotateNhoods(milo, da_results, coldata_col = "seurat_clusters")

```


### CNA

```{r}
disease_group = "All"
disease_label = "AllSS"
control_group = c("nonSS")

logi = (gex$Disease %in% c(control_group,
                           disease_group
))
gex_sub <- subset(x = gex, 
                  Disease %in% control_group | 
                    Disease %in% disease_group
)

gex_sub = FindNeighbors(gex_sub, reduction = "harmony", 
                        dims = 1:dim,
                        k.param = 30,
                        verbose = FALSE
)

gex_sub@meta.data$Disease <- as.character(gex_sub@meta.data$Disease)
gex_sub@meta.data$condition <- ifelse(gex_sub@meta.data$Disease %in% control_group, "nonSS", "SS")
gex_sub@meta.data$condition_val <- as.numeric(factor(gex_sub@meta.data$condition, c('nonSS', 'SS')))
print(table(gex_sub@meta.data$condition))
print(table(gex_sub@meta.data$Sex))

covs = c("Age","Sex")

gex_sub <- association.Seurat(
  seurat_object = gex_sub, 
  test_var = 'condition_val', 
  samplem_key = 'orig.ident', 
  graph_use = 'RNA_snn', 
  verbose = TRUE,
  batches = NULL,
  covs = covs
)


```

<br/><br/>

***

# sessionInfo

```{r}
sessionInfo()
```
